---
title: "5300 Predictive Analysis Competition"
author: "Vivian Wen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="", warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=55), tidy = TRUE)
```

## Database setup:

Read database 'songs' and 'scoring data' from directory:

```{r}
songs = read.csv('C:/Users/vivia/OneDrive/桌面/APAN5200/analysisData.csv')
scoringData = read.csv('C:/Users/vivia/OneDrive/桌面/APAN5200/scoringData.csv')
```

The initial model given by professor:
```{r}
model = lm(rating~ tempo+time_signature,songs)
pred = predict(model,newdata=scoringData)
```


## Unsuccessful Models I have tried:

The linear regression model:
```{r}
model1 = lm(rating~time_signature+tempo+track_duration+danceability+acousticness+valence+loudness+speechiness,songs)
summary(model1)
pred1 = predict(model1,newdata=scoringData)
```

regression models are not good without solve for Multicollinearity and heteroskedasticity.

However, the variables do contain some correlations between variables.

```{r}
cor(songs$rating, songs$danceability) #0.1388896
cor(songs$rating, songs$energy) #0.1024138
cor(songs$rating, songs$time_signature) #0.09239721
cor(songs$rating, songs$tempo) #0.0126714
cor(songs$rating, songs$valence) #-0.09392114
cor(songs$rating, songs$track_duration) #0.1467323
cor(songs$rating, songs$track_explicit) #0.1454988
cor(songs$track_duration, songs$track_explicit) #0.03321112
cor(songs$rating, songs$instrumentalness) #-0.08836823
cor(songs$rating, songs$energy) #0.1024138
cor(songs$rating, songs$key) #0.001075925
cor(songs$rating, songs$loudness) #0.1964508
cor(songs$rating, songs$speechiness) #0.07675559
cor(songs$rating, songs$acousticness) #-0.1978989
cor(songs$rating, songs$liveness) #-0.05878034
cor(songs$energy, songs$loudness) #0.6807528
cor(songs$time_signature, songs$tempo) #-0.01463417
cor(songs$danceability, songs$acousticness) #-0.3140044
cor(songs$liveness, songs$loudness) #0.04015022
```

The Fitting Generalized Linear Model:
```{r}
model2 = glm(rating~danceability+acousticness+valence+loudness, data = songs)
summary(model2)
pred2 = predict(model2,newdata=scoringData)
```

Use generalized additive model:
```{r}
library(mgcv)
model4 = gam(rating~time_signature+energy+tempo+track_duration+track_explicit+danceability+s(acousticness)+s(instrumentalness)+valence+loudness+energy+liveness, method = 'REML', data = songs)
summary(model4)
pred4 = predict(model4,newdata=scoringData)
```


Use variables with random forest(without data tidying):
```{r}
library(randomForest)
#library(tidyverse)
set.seed(1031)
model3 = randomForest(rating~time_signature+energy+mode+tempo+track_duration+track_explicit+danceability+acousticness+valence+loudness+speechiness+liveness+instrumentalness+key,data = songs, importance = TRUE)
pred3 = predict(model3,newdata=scoringData)
```


Use rpart model(but the result is not good compare to random forest:
```{r}
library(rpart)

tree1 = rpart(rating~time_signature+energy+mode+tempo+track_duration+track_explicit+danceability+acousticness+valence+loudness+speechiness+liveness+instrumentalness+key,data = songs, method = 'anova', control= rpart.control(cp=0.0001))
pred_tree = predict(tree1,newdata=scoringData, type='vector')
```

Try forest ranger:
```{r}
library(ranger)
library(caret)

set.seed(1031)
forest_ranger = ranger(rating~time_signature+energy+mode+tempo+track_duration+track_explicit+danceability+acousticness+valence+loudness+speechiness+liveness+instrumentalness+key,
                       data = songs, 
                       num.trees = 1000)
pred_train = predict(forest_ranger, data=scoringData, num.trees = 1000)
#15.15955

```

Tried tuned ranger but didn't run a result:
```{r}
trControl=trainControl(method="cv",number=5)
tuneGrid = expand.grid(mtry= 1:10, 
                       splitrule = c('variance','extratrees','maxstat'), 
                       min.node.size = c(2,3,4,5,10))
set.seed(1031)
cvModel = train(rating~time_signature+mode+tempo+track_duration+track_explicit+danceability+acousticness+valence+loudness+speechiness+liveness+instrumentalness+key,
                data=songs,
                method="ranger",
                num.trees=1000,
                trControl=trControl,
                tuneGrid=tuneGrid)
cvModel$bestTune
```


```{r}
set.seed(1031)
cv_forest_ranger = ranger(rating~time_signature+mode+tempo+track_duration+track_explicit+danceability+acousticness+valence+loudness+speechiness+liveness+instrumentalness+key,
                          data=songs,
                          num.trees = 1000, 
                          mtry=cvModel$bestTune$mtry, 
                          min.node.size = cvModel$bestTune$min.node.size, 
                          splitrule = cvModel$bestTune$splitrule)

pred_forest_ranger = predict(cv_forest_ranger, data = scoringData, num.trees = 1000)
#15.20061
```




Then, I found out it would be better if I can do some data tidying and seperate the genre column to dummy variables.
So, I began to do some data tidying with genre column:

### Step1: Data Tidying: clean and managing the genre column into dummy variables and to various columns

1. clean the format of genre column from both songs and scoringData datasets, and make them into matrix.

```{r}
songs$genre <- gsub("\\[|\\]", "", songs$genre)
songs$genre <- gsub("'","",songs$genre)
songs$genre <- gsub("-"," ",songs$genre)

x  <- strsplit(songs$genre, ",")
x1 <- unique(unlist(x))
genres <- t(sapply(x,function(z)table(factor(z,levels=x1))))


scoringData$genre <- gsub("\\[|\\]", "", scoringData$genre)
scoringData$genre <- gsub("'","",scoringData$genre)
scoringData$genre <- gsub("-"," ",scoringData$genre)

a  <- strsplit(scoringData$genre, ",")
a1 <- unique(unlist(a))
genres_test <- t(sapply(a,function(z)table(factor(z,levels=a1))))

```

2. Merge the dummy variables into dataframe. Note that I did not use all genre types but I chose to only use count greater than and equal to 5.
```{r}
library(dplyr)
library(tidyr)
library(janitor)

colnames(genres) <- trimws(colnames(genres), whitespace = "[\\h\\v]")
colnames(genres) <- gsub("\"","",colnames(genres))
genres = as.data.frame(genres)

merged_genres <- as.data.frame(t(apply(genres,1, function(x) tapply(x,colnames(genres),sum))))
colnames(merged_genres) <- gsub(" ","_",colnames(merged_genres))
merged_genres <- merged_genres %>%
  clean_names()

sum_genre <- colSums(merged_genres)

# for scoring Data:
colnames(genres_test) <- trimws(colnames(genres_test), whitespace = "[\\h\\v]")
colnames(genres_test) <- gsub("\"","",colnames(genres_test))
genres_test = as.data.frame(genres_test)

merged_genres_test <- as.data.frame(t(apply(genres_test,1, function(x) tapply(x,colnames(genres_test),sum))))
colnames(merged_genres_test) <- gsub(" ","_",colnames(merged_genres_test))
merged_genres_test <- merged_genres_test %>%
  clean_names()


colname1 <- colnames(merged_genres)
colname2 <- colnames(merged_genres_test)
common_col <- intersect(colname1, colname2)


good_genres <- merged_genres[, common_col]
colnames(good_genres) <- gsub(" ", "_", colnames(good_genres))

good_genres_test <- merged_genres_test[, common_col]
colnames(good_genres_test) <- gsub(" ", "_", colnames(good_genres_test))

genres_greater20 <- good_genres[, colSums(good_genres) >= 5]
genres_test_greater20 <- good_genres_test[, colSums(good_genres_test) >= 5]


genreName1 <- colnames(genres_greater20)
genreName2 <- colnames(genres_test_greater20)
intersectionCol <- intersect(genreName1, genreName2)

top_genre1 <- genres_greater20[, intersectionCol]
top_genre2 <- genres_test_greater20[, intersectionCol]

```


### The model I ran after data tidying with the genre column only:

```{r}
library(ranger)
set.seed(1031)
forest_ranger4 = ranger(selected_formula,
                       data = songs_sum, 
                       num.trees = 1000)
pred_forest_ranger4 = predict(forest_ranger4, data=scoringData_sum, num.trees = 1000)
```


## Later, I found out there could be better perdictions when I did both data tidying for genre and performer columns together:
### Data tidying with performer column:

```{r}
songs$performer <- gsub("'","",songs$performer)
songs$performer <- gsub(" ","_",songs$performer)
songs$performer <- gsub("\\(|\\)","",songs$performer)
songs$performer <- gsub("\\[|\\]", "", songs$performer)


c = strsplit(songs$performer,'_Featuring_|_featuring_|_,_|_with_|_With_|_And_|_&_|&|,_')
c1 <- unique(unlist(c))
performer <- t(sapply(c,function(z)table(factor(z,levels=c1))))
performer <- as.data.frame(performer)
sum_singer = rowSums(performer)

#For "Scoring Data":
scoringData$performer <- gsub("'","",scoringData$performer)
scoringData$performer <- gsub("'", "_", scoringData$performer)
scoringData$performer <- gsub("\\(|\\)","",scoringData$performer)
scoringData$performer <- gsub("\\[|\\]", "", scoringData$performer)

d = strsplit(scoringData$performer,'_Featuring_|_featuring_|_,_|_with_|_With_|_And_|_&_|&|,_')
d1 <- unique(unlist(d))
performer_test <- t(sapply(d,function(z)table(factor(z,levels=d1))))


colnames(performer_test) <- trimws(colnames(performer_test), whitespace = "[\\h\\v]")
colnames(performer_test) <- gsub("\"","",colnames(performer_test))
performer_test = as.data.frame(performer_test)

library(dplyr)
library(tidyr)
library(janitor)
performer <- performer %>%
  clean_names()

performer_test <- performer_test %>%
  clean_names()

colname1p <- colnames(performer)
colname2p <- colnames(performer_test)
common_colp <- intersect(colname1p, colname2p)


good_performer <- performer[, common_colp]
colnames(good_performer) <- gsub(" ", "_", colnames(good_performer))

good_performer_test <- performer_test[, common_colp]
colnames(good_performer_test) <- gsub(" ", "_", colnames(good_performer_test))

df1 <- performer[, common_colp]
df2 <- performer_test[, common_colp]

top_df1 <- good_performer[, colSums(df1) > 5]
top_df2 <- good_performer_test[, colSums(df2) > 5]
top <- colnames(top_df1)
top2 <- colnames(top_df2)
top_intersact <- intersect(top,top2)

top_perform1 <- top_df1[, top_intersact]
top_perform2 <- top_df2[, top_intersact]

```

```{r}
songs_sum <- cbind.data.frame(songs, top_genre1, top_perform1)

songs_sum <- songs_sum %>%
  relocate(rating, .after = last_col())

scoringData_sum <- cbind.data.frame(scoringData, top_genre2, top_perform2)
```


### Formula for predictive model
```{r}
name <- c('track_duration', 'track_explicit', 'danceability', 'energy',
                'key','loudness', 'mode', 'speechiness', 'acousticness',
                'instrumentalness', 'liveness', 'valence',
                'tempo', 'time_signature', intersectionCol, top_intersact)
Formula <- formula(paste("rating~", 
     paste(name, collapse="+")))

```

### I tried to do feature selction with my forest ranger model, but the results is not good as my model without feature selection.
```{r}
library(randomForestSRC)
 rf <- rfsrc(Formula, data = songs_sum)
important <- vimp(rf)$importance
name_important <- names(important)

Formula2 <- formula(paste("rating~", 
     paste(name_important, collapse="+")))


library(ranger)
library(caret)
library(ggplot2)

set.seed(1031)
forest_ranger = ranger(Formula2,
                       data = songs_sum, 
                       num.trees = 12000)

pred_forest_ranger = predict(forest_ranger, data=scoringData_sum, num.trees = 12000)

```


## The best model with lowest RMSE in Kaggle
```{r}
#Forest Ranger with new model:forest_ranger2

library(ranger)
library(caret)
library(ggplot2)

set.seed(1031)
forest_ranger = ranger(Formula,
                       data = songs_sum, 
                       num.trees = 12000)

pred_forest_ranger = predict(forest_ranger, data=scoringData_sum, num.trees = 12000)

```


## The RMSE calculation formula I have used:
```{r}
rmse = sqrt(mean(pred_forest_ranger$predictions-songs_sum$rating)^2)
rmse
```

## Create output with predictions:
```{r}
submissionFile = data.frame(id = scoringData$id, rating = pred_forest_ranger$predictions)
write.csv(submissionFile, 'C:/Users/vivia/OneDrive/桌面/APAN5200/submissionFile_Vivian.csv',row.names = F)
```
